LFQ: Online Learning of Per-flow Queuing Policies using Deep Reinforcement Learning

### 2020-07-03 paper/paper.tex
Add neural network weights and new paper version

\begin{abstract}
The increasing number of different, incompatible congestion control algorithms has led to an increased deployment of fair queuing. Fair queuing isolates each network flow and can thus guarantee fairness for each flow even if the flows' congestion controls are not inherently fair. So far, each queue in the fair queuing system either has a fixed, static maximum size or is managed by an Active Queue Management (AQM) algorithm like CoDel. In this paper we design an AQM mechanism (Fair Learning Qdisc (LFQ)) that dynamically learns the optimal buffer size for each flow according to a specified utility function online. We show that our Deep Learning based algorithm can dynamically assign the optimal queue size to each flow depending on its congestion control, delay and bandwidth. Comparing to competing fair AQM schedulers, it provides significantly smaller queues while achieving the same or higher throughput. 
\end{abstract}

@@

New \glspl{cca} for TCP and QUIC are still being introduced revisiting old ideas but also introducing new concepts \cite{dong_pcc_2018, cardwell_bbr:_2016, hock_tcp_2017, bachl_cocoa_2019, jay_deep_2019}. Some of them, such as BBR, are already being used in real-world production systems. While new \glspl{cca} are commonly being designed with compatibility to other \glspl{cca} in mind, often they do not share the link completely fairly with older \glspl{cca} \cite{hock_experimental_2017, fejes_who_2019, fejes_incompatibility_2020}. Besides that, network flows using the same \gls{cca} can also be unfair to each other: For example, BBR favors flows with a high \gls{rtt}, while New Reno favors those with a low one \cite{turkovic_interactions_2019,turkovic_fifty_2019}. This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.

Given the high throughput and low queuing delay that is required by emerging cloud gaming \cite{jarschel_evaluation_2011} and virtual reality \cite{elbamby_toward_2018} applications, the optimal management of each flow's queue becomes an increasingly relevant question. Furthermore, for 5G, ultra low latency communications have been made a priority \cite{li_5g_2018} and we argue that effective buffering must also be taken into account for achieving this goal. The simplest solution to managing a flow's buffer is to a static buffer size (queue size) for each flow. However, this can lead to inacceptably large queuing delays. Another solution is to use an advanced \gls{aqm} mechanism like CoDel for each flow. CoDel aims to keep the queue length under a certain threshold, making sure that the queuing delay was smaller than 5\,ms in the last 100\,ms at least once. Otherwise it drops packets to decrease the queue length. Both of these approaches do not differentiate between flows: They apply the same logic to each flow irrespective of its congestion control and irrespective of the current bandwidth and \gls{rtt}. \cite{bachl_cocoa_2019} showed that this behavior leads to some flows not being able to claim the full bandwidth that they are entitled to. Other flows might achieve the full bandwidth but keep an unnecessary standing queue. The authors then proposed a mechanism that adjusts the queue of each flow based on its congestion control and showed that it works well for several common \glspl{cca}. However, their algorithm has parameters that have to be manually adjusted and it is not guaranteed to work for every \gls{cca} because it makes the assumption that each flow's congestion window follows a zigzag pattern. Furthermore, their approach is only tailored towards flows that always have data to send while the behavior for application limited flows, that also have idle periods, is not clear.

Instead of a hand-crafted solution like the existing ones, we argue that instead operators of network hardware should be able to simply specify a utility function, and the \gls{aqm} should then automatically find the right queuing policy based on that utility function using \gls{ml}.

@@

\cite{kim_deep_2019} use Deep Reinforcement Learning to develop an \gls{aqm} mechanism for IoT devices that works on the output queue of a switch.  \cite{bouacida_practical_2019,bisoy_design_2017} use \gls{rl} to build an \gls{aqm} algorithm. \cite{vucevic_reinforcement_2007} uses \gls{rl} as well and focuses specifically on cellular communications. \cite{lin_kemy_2015} use a custom optimization approach that relies on as offline greedy algorithm which finds optimal rules for an \gls{aqm} that fulfills a certain utility function. \cite{shah_sam_2016} build an \gls{aqm} mechanism, which utilizes support vector machines to learn an optimal queuing policy for a shared buffer.

@@

If the buffer for a flow is too small, this flow cannot achieve the full bandwidth as shown in \autoref{fig:tooLittle}. On the other side, if the buffer is too large (\autoref{fig:tooMuch}), the flow can achieve the full bandwidth but a standing queue exists, which causes unnecessary delay.

For the popular \gls{cca} \textit{New Reno}, the buffer size that is necessary in a fair queuing setting is one \gls{bdp}, meaning $\textit{speed}\times\textit{delay}$ (\autoref{fig:perfect}). This is because its multiplicative decrease factor is $0.5$, meaning that after packet loss, the congestion window is halved. For other \glspl{cca} like \textit{Cubic} \cite{ha_cubic:_2008}, the multiplicative decrease factor is $0.7$, meaning that the congestion window is reduced to 70\% of its previous value upon packet loss. This means that the minimum buffer size required to achieve full throughput for a Cubic flow is $\left(\frac{1}{0.7}-1\right)\times \textit{BDP}$, which is around 43\% of the \gls{bdp} and thus less than for New Reno. That means the minimum required buffer size depends on the \gls{cca}. To reduce packet loss and standing queues we propose an adaptive fair \gls{aqm} that learns the available bandwidth, \gls{rtt} and the underlying \gls{cca} of a flow to adjust the buffer optimally.

@@

To learn an optimal \gls{aqm} policy, we specify a utility function, which the \gls{ml} based \gls{aqm} should learn to optimize for each flow in a \gls{rl} fashion. A simple and effective utility function is for example:

\begin{align}
\textit{Utility} = \textit{bandwidth}-\alpha\times\textit{queue size}
\label{eq:utility}
\end{align}

In this utility function, the choosable parameter $\alpha$ specifies the tradeoff that is chosen between the bandwidth and the buffer size. With $\alpha$ going to zero, the optimum policy approaches the one in which the buffer size is large enough that a flow never underutilizes the link but at the same time the buffer is never going to be larger than necessary if this doesn't provide a benefit in throughput. Alternatively, the delay could be used instead of the queue size because it is more closely related to \gls{qoe}. However, the delay that is caused by a certain queue length depends on the current link speed. This means that using delay would make the utility function more variable, depending on the current link speed.

Our \gls{ml} system uses the above utility function to learn the optimal behavior. \gls{ours} computes features for each flow after every packet it receives and then outputs the optimal buffer size of which it thinks that it maximizes the utility function. As the input we use the following features, which are updated continuously when receiving/sending a packet: 
\begin{itemize}
\item queue size
\item standard deviation of the queue size 
\item maximimum allowed buffer size
\item rate of incoming data
\item rate of outgoing data
\item time since the last packet loss
\end{itemize}
We do not use these features directly because this could result in large, spontaneous variation in the input features. For example, two packets might not be received in regular intervals but in bursts. This means that without any smoothening, large instability would occur. Instead use 10 exponentially weighted averages of them for each feature with weights of $2^{-4}$, $2^{-5}$, ... , $2^{-13}$. The advantage of using exponentially weighted averages is that they do not occupy any space in memory except for their own values. This is opposed to regular moving averages, which have to keep the entire window of data in memory. For example, for a regular weighted average of window size 1000, the last 1000 data points have to be kept in memory at any time, leading to a large consumption of memory.

One difficulty is to get an exponentially moving average for the rate of incoming/outgoing packets. Instead we compute the exponentially moving average of the interarrival times of incoming packets and the interdeparture times of outgoing packets. We then invert this number to get the rate. An issue with this approach is that numeric instability and division by zero errors can occur using this approach. Thus, we only use the exponential average with a weight of $2^{-n}$ after at least $n$ packets have arrived. Otherwise we set it to zero. We did this because we saw that using the averages of interarrival/interdeparture times led to slower learning than using the rates. This is because if only the interarrival/interdeparture times are given, the neural network also has to learn the additional step of inverting them, making the learning more time-consuming.

@@

For the simulations we randomly draw a bandwidth (5 to 25\,Mbit/s), a delay (5 to 25\,ms) and duration (3.75\,ms to 6.25\,ms) and a \gls{cca} (New Reno or BIC \cite{lisong_xu_binary_2004}). We chose these ranges bandwidths and delays so that the neural network encounters flows with vastly different characteristics and has to learn to deal with them. We use New Reno because it is one of the oldest, most widely deployed \glspl{cca} and BIC because it is similar to Cubic, which is the default in Linux, Windows and most other OSs currently. We cannot use Cubic itself because there is no stable Cubic implementation in ns-3 currently.

As the experiment time (the time at which the experiment is launched to determine if a larger or smaller buffer size would be better), we use a random number between 0 and half of the flow duration of the current flow. The number must be random so that the \gls{rl} learn to take decisions for every stage of a flow: The optimal maximum buffer size is different in the beginning of the flow, when you don't know any information about the flow yet, than it is after several seconds. The expected behavior is that in the beginning of a flow, the \gls{rl} simply uses the average buffer size that works the best on average for all flows. Then, after more information about the flow is known and it can be fingerprinted effectively, the estimation of the \gls{rl} is supposed to become more fine-grained and specific to the current flow. That means that the \gls{rl} has to learn what the optimal buffer size is for every stage of a flow and that's why the time at which the experiment is performed must be random. Another issue is that if, for example, the experiment is performed just before the end of the flow, the impact of the decision of the \gls{rl} is not clear: If the \gls{rl}, for example, outputs a maximum buffer size of 15 just before the end of the flow, then it has no influence on the flow because the flow ends anyway. That's why we don't perform experiments for the \gls{rl} training in the latter half of a flow.

For the simulation we use two hosts that are directly connected to each other. We also experimented initially with a switch connecting these hosts but this made the simulation significantly slower. Thus we made the simulation as simple as possible and implemented our \gls{aqm} in front of the bottleneck link, that connects the sender to the receiver. For all other buffers we use a FIFO queue with a buffer of 100 packets (the default).

For the \gls{dl}, we used a fully connected neural network consisting of an input layer, three layers of 256 neurons which each have the leaky ReLU \cite{noauthor_rectifier_2020} function applied and finally an output layer which has an output size of 1. All neural networks we use have this architecture. For the learning we chose gradient descent with a learning rate of $0.01$. To save computational resources, how often our \gls{dl} logic runs and outputs the optimal buffer size is a configurable parameter. The most fine-grained would be to let the \gls{dl} run for each packet that is received. However, this would be a computational waste and thus, by default, we only let the \gls{dl} run for every 10th packet. Other sampling intervals can be configured. Also, an option would be to not let the \gls{dl} run every $n$th packet but every $x$ milliseconds.

@@

The training procedure involves the following steps:
\begin{enumerate}
\item Draw a random sample of 20 bandwidths, delays, \gls{cca} and flow durations.
\item Simulate each flow concurrently, compute the feature vector continuously and let the neural network output its optimal buffer size each time a new packet arrives. 
\item During each flow, at a random time between 0 and half the flow length in seconds, perform the experiment: Fork the simulation process for each flow and continue one simulation with the current buffer size $+1$ packet and one with the current buffer size $-1$ packet until the end of the flow. This procedure is essentially \textit{A/B testing}.  
\item Wait until all flows are finished
\item Check for each flow which of the two versions performed better ($+1$ or $-1$) regarding the utility function.
\item Update the neural network so that it learns to output the better buffer size when being fed the inputs as they were at the time at which the experimentation started. Specifically, we compute the \gls{mae} between the output of the neural network and the desired output (the one that performed better in the experiment) for the whole batch of 20 results.
\item Start the next iteration. 
\end{enumerate}

@@

\begin{figure*}[h]
\centering
\subfloat[New Reno, varying bandwidth.\label{fig:SmallAlphaNewRenoBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_New Reno_bandwidth"}.pdf}
}{}
\subfloat[New Reno, varying delay.\label{fig:SmallAlphaNewRenoDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_New Reno_delay"}.pdf}
}{}
\subfloat[BIC, varying bandwidth\label{fig:SmallAlphaBicBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_Bic_bandwidth"}.pdf}
}{}
\subfloat[BIC, varying delay\label{fig:SmallAlphaBicDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_Bic_delay"}.pdf}
}{}
\caption{Change of the buffer size after offline learning when varying bandwidth/delay/\gls{cca}. When bandwidth is varied, delay is kept in the middle of its range, at 15\,ms. Vice versa, when delay is varied, bandwidth is kept in the middle of its range, at 15\,Mbit/s.}
\label{fig:offlineSmallAlpha}
\end{figure*}

When performing offline learning with the tradeoff parameter $\alpha=0.01$ (\autoref{fig:offlineTraining}), the neural network first starts with a prediction that is close to 0 (due to the initialization of neural network weights). During the first couple of thousand training flows, it increases its output and approaches what appears to be the buffer size that works best on average (slightly higher than 15). Then the neural network gradually starts understanding and differentiating different flows and learns custom policies for them. After around 250000 flows, the output appears to be stable and doesn't change anymore. There's one noticeable outlier at around 120000 flows. This appears to be some instability that can commonly occur in the early stages of neural network training but what has no significance for the training process since it resumes normally after the outlier.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_0_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoLargeBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoSmallBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_1_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a BIC flow controlled by \gls{ours} ($\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleBicLargeBw}
\end{figure}

@@

\autoref{tab:avgOfflineSmallAlpha} shows that the average output maximum queue size is more than double for New Reno than for BIC. This shows that the neural network also learned to distinguish between these two \glspl{cca}. Since there are no features that directly indicate the \gls{cca}, the neural network must have learned these by combining the other features and observing distinct behavior for New Reno and BIC. This is expected, since BIC has a different multiplicative decrease factor than New Reno ($0.8$ vs $0.5$) the neural network should learn to output different maximum queue sizes for these two. Specifically, we would expect BIC to need a smaller buffer as its decrease is smaller, thus requiring a smaller buffer.

@@

\autoref{fig:offlineTrainingLargeAlpha} shows that also for the larger $\alpha$, the neural network outputs converge quickly after a couple of 10000 flows.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-4-15-39-25_108000.weights_cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} ($\alpha=10$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue occasionally becomes empty which is intended as with the higher tradeoff $\alpha$, \gls{ours} tries to favor a small queue more than high throughput.}
\label{fig:exampleRenoSmallBwLargeAlpha}
\end{figure}

@@

Results in \autoref{tab:corrOfflineLargeAlpha} and \autoref{tab:corrOnlineLargeAlpha} show that for a larger $\alpha$ \gls{ours} still works as expected and that the average output maximum queue size is smaller, as expected. p

@@

In contrast to offline learning online learning enables to train in deployment. This makes it possible to adapt the behavior slowly over time (over a time span of month or years), for example if new \glspl{cca} emerge. Furthermore, training with real flows is more realistic than training only with simulated flows in a simulation.

@@

We envision \gls{ours} being used on switches and routers close to the Internet's edge, as here usually the bottleneck links are and because network devices here only have to manage a rather small number of flows compared to the Internet's core. An interesting future direction of work could be to envision different utility functions, for example directly based on \gls{qoe}.

>>>>

\begin{abstract}
The increasing number of different, incompatible congestion control algorithms has led to an increased deployment of fair queuing. Fair queuing isolates each network flow and can thus guarantee fairness for each flow even if the flows' congestion controls are not inherently fair. So far, each queue in the fair queuing system either has a fixed, static maximum size or is managed by an Active Queue Management (AQM) algorithm like CoDel. In this paper we design an AQM mechanism (Fair Learning Qdisc (LFQ)) that dynamically learns the optimal buffer size for each flow according to a specified reward function online. We show that our Deep Learning based algorithm can dynamically assign the optimal queue size to each flow depending on its congestion control, delay and bandwidth. Comparing to competing fair AQM schedulers, it provides significantly smaller queues while achieving the same or higher throughput. 
\end{abstract}

@@

New \glspl{cca} for TCP and QUIC are still being introduced, revisiting old ideas but also introducing new concepts \cite{dong_pcc_2018, cardwell_bbr:_2016, hock_tcp_2017, bachl_cocoa_2019, jay_deep_2019}. Some of them, such as BBR, are already being used in real-world production systems. While new \glspl{cca} are commonly being designed with compatibility to other \glspl{cca} in mind, often they do not share the link completely fairly with older \glspl{cca} \cite{hock_experimental_2017, fejes_who_2019, fejes_incompatibility_2020}. Besides that, network flows using the same \gls{cca} can also be unfair to each other: For example, BBR favors flows with a high \gls{rtt}, while New Reno favors those with a low one \cite{turkovic_interactions_2019,turkovic_fifty_2019}. This unfairness can be mitigated by using \gls{fq} at the bottleneck link, isolating each flow from all other flows and assigning each flow an equal share of bandwidth \cite{dumazet_pkt_sched:_2013}.

Given the high throughput and low queuing delay that is required by emerging cloud gaming \cite{jarschel_evaluation_2011} and virtual reality \cite{elbamby_toward_2018} applications, the optimal management of each flow's queue becomes an increasingly relevant question. Furthermore, for 5G, ultra low latency communications have been made a priority \cite{li_5g_2018} and we argue that effective buffering must also be taken into account for achieving this goal. The simplest solution manage a flow's buffer is to use a static buffer size (queue size) for each flow. However, this can lead to inacceptably large queuing delays. Another solution is to use an advanced \gls{aqm} mechanism like CoDel for each flow. CoDel aims to keep the queue length under a certain threshold, making sure that the queuing delay was smaller than 5\,ms in the last 100\,ms at least once. Otherwise it drops packets to decrease the queue length. Both of these approaches do not differentiate between flows: They apply the same logic to each flow irrespective of its congestion control and irrespective of the current bandwidth and \gls{rtt}. \cite{bachl_cocoa_2019} showed that this behavior leads to some flows not being able to claim the full bandwidth that they are entitled to. Other flows might achieve the full bandwidth but keep an unnecessary standing queue. The authors then proposed a mechanism that adjusts the queue of each flow based on its congestion control and showed that it works well for several common \glspl{cca}. However, their algorithm has parameters that have to be manually adjusted and it is not guaranteed to work for every \gls{cca} because it makes the assumption that each flow's congestion window follows a zigzag (sawtooth) pattern. Furthermore, their approach is only tailored towards flows that always have data to send while the behavior for application limited flows, that also have idle periods, is not clear.

Instead of a hand-crafted solution like the existing ones, we argue that instead operators of network hardware should be able to simply specify a reward function, and the \gls{aqm} should then automatically find the right queuing policy based on that reward function using \gls{ml}. We propose \gls{ours}, which achieves these goals.

@@

\cite{kim_deep_2019} use Deep Reinforcement Learning to develop an \gls{aqm} mechanism for IoT devices that works on the output queue of a switch.  \cite{bouacida_practical_2019,bisoy_design_2017} use \gls{rl} to build an \gls{aqm} algorithm for shared buffers. \cite{vucevic_reinforcement_2007} uses \gls{rl} as well and focuses specifically on cellular communications. \cite{lin_kemy_2015} use a custom optimization approach that relies on as offline greedy algorithm which finds optimal rules for an \gls{aqm} that fulfills a certain reward function. \cite{shah_sam_2016} build an \gls{aqm} mechanism which utilizes support vector machines to learn an optimal queuing policy for a shared buffer.

@@

If the buffer for a flow is too small, a flow cannot achieve the full bandwidth as shown in \autoref{fig:tooLittle}. On the other side, if the buffer is too large (\autoref{fig:tooMuch}), the flow can achieve the full bandwidth but a standing queue exists, which causes unnecessary delay.

For the popular \gls{cca} \textit{New Reno}, the buffer size that is necessary in a fair queuing setting is one \gls{bdp}, meaning $\textit{speed}\times\textit{delay}$ (\autoref{fig:perfect}). This is because its multiplicative decrease factor is $0.5$, meaning that after packet loss, the congestion window is halved. For other \glspl{cca} like \textit{Cubic} \cite{ha_cubic:_2008}, the multiplicative decrease factor is $0.7$, meaning that the congestion window is reduced to 70\% of its previous value upon packet loss. This means that the minimum buffer size required to achieve full throughput for a Cubic flow is $\left(\frac{1}{0.7}-1\right)\times \textit{BDP}$, which is around 43\% of the \gls{bdp} and thus less than for New Reno. That means the minimum required buffer size, among other factors, depends on the \gls{cca} and specifically the \gls{cca}'s multiplicative decrease factor. To reduce packet loss and standing queues we propose an adaptive fair \gls{aqm} that learns the available bandwidth, \gls{rtt} and the underlying \gls{cca} of a flow to adjust the buffer optimally.

@@

To learn an optimal \gls{aqm} policy, we specify a reward function, which the \gls{ml} based \gls{aqm} should learn to optimize for each flow in a \gls{rl} fashion. A simple and effective reward function is for example:

\begin{align}
\textit{Reward} = \textit{bandwidth}-\alpha\times\textit{queue size}
\label{eq:reward}
\end{align}

In this reward function, the choosable parameter $\alpha$ specifies the tradeoff that is chosen between the bandwidth and the buffer size. With $\alpha$ going to zero, the optimum policy approaches the one in which the buffer size is large enough that a flow never underutilizes the link but at the same time the buffer is never going to be larger than necessary if this doesn't provide a benefit in throughput. Alternatively, the delay could be used instead of the queue size because it is more closely related to \gls{qoe}. However, the delay that is caused by a certain queue length depends on the current link speed. This means that using delay would make the reward function more variable, depending on the current link speed.

Our \gls{ml} system uses the above reward function to learn the optimal behavior. \gls{ours} outputs the optimal buffer size of which it thinks that it maximizes the reward function when receiving a packet. As the input for this decision we use the following features, which are updated continuously when receiving/sending a packet: 
\begin{itemize}
\item queue size
\item standard deviation of the queue size 
\item maximimum buffer size
\item rate of incoming data
\item rate of outgoing data
\item time since the last packet loss
\end{itemize}
We do not use these features directly because this could result in large, spontaneous variation in the input features. For example, packets might not be received in regular intervals but in bursts. This means that without any smoothening, large instability would occur. Thus, instead we use 10 exponentially weighted averages of each feature with weights of $2^{-4}$, $2^{-5}$, ... , $2^{-13}$. The advantage of using exponentially weighted averages is that they do not occupy any space in memory except for their own values. This is opposed to regular moving averages, which have to keep the entire window of data in memory. For example, for a regular weighted average of window size 1000, the last 1000 data points have to be kept in memory at any time, leading to a large consumption of memory.

One difficulty is to get an exponentially moving average for the rate of incoming/outgoing packets. Instead we compute the exponentially moving average of the interarrival times of incoming packets and the interdeparture times of outgoing packets. We then invert this number to get the rate. An issue with this approach is that numeric instability and division by zero errors can occur using this approach. Thus, we only use the exponential average with a weight of $2^{-n}$ after at least $n$ packets have arrived. Otherwise we set it to zero. We did this inversion because we saw that using the averages of interarrival/interdeparture times led to slower learning than using the rates. This is because if only the interarrival/interdeparture times are given, the neural network also has to learn the additional step of inverting them, making the learning more time-consuming.

@@

For the simulations we randomly draw a bandwidth (5 to 25\,Mbit/s), a delay (5 to 25\,ms) and duration (3.75\,ms to 6.25\,ms) and a \gls{cca} (New Reno or BIC \cite{lisong_xu_binary_2004}). We chose these ranges of bandwidths and delays so that the neural network encounters flows with vastly different characteristics and has to learn to deal with them. We use New Reno because it is one of the oldest, most widely deployed \glspl{cca} and BIC because it is similar to Cubic, which is the default in Linux, Windows and most other OSs currently. We cannot use Cubic itself because there is no stable Cubic implementation in ns-3 currently.

During simulating each flow, at a certain point, we perform an experiment to let the \gls{rl} learn what the optimal buffer size in the current scenario for the current flow would be. As the experiment time, we use a random number between 0 and half of the flow duration of the current flow. The number must be random so that the \gls{rl} learns to take decisions for every stage of a flow: The optimal maximum buffer size is different in the beginning of the flow, when any information about the flow is known yet, than it is after several seconds. The expected behavior is that in the beginning of a flow, the \gls{rl} simply uses the average buffer size that works the best on average for all flows. Then, after more information about the flow is known and it can be fingerprinted effectively, the estimation of the \gls{rl} is supposed to become more fine-grained and specific to the current flow. That means that the \gls{rl} has to learn what the optimal buffer size is for every stage of a flow and that's why the time at which the experiment is performed must be random. Another issue is that if, for example, the experiment is performed just before the end of the flow, the impact of the decision of the \gls{rl} is not clear: If the \gls{rl}, for example, outputs a maximum buffer size of 15 just before the end of the flow, then it has no influence on the flow because the flow ends anyway. That's why we don't perform experiments for the \gls{rl} training in the latter half of a flow.

For the simulations we use two hosts that are directly connected to each other. We also experimented initially with a switch connecting these hosts but this made the simulation significantly slower. Thus we made the simulation as simple as possible and implemented our \gls{aqm} in front of the bottleneck link that connects the sender to the receiver. For all other buffers we use a FIFO queue with a buffer of 100 packets (the default).

For the \gls{dl}, we used a fully connected neural network consisting of an input layer, three layers of 256 neurons which each have the leaky ReLU \cite{noauthor_rectifier_2020} function applied and finally an output layer which has an output size of 1. All neural networks we use have this architecture. We chose this number of neurons and layers as recent research shows that generally, overfitting is not a problem, if a neural network is trained long enough \cite{nakkiran_deep_2019}. Thus, it is generally not bad to have more neurons and layers than necessary. That's why we opted for 256 neurons and 3 layers, which we think is quite generous considering that there are only 60 input features and one output. For the learning we chose gradient descent with a learning rate of $0.01$. To save computational resources, how often our \gls{dl} logic runs and outputs the optimal buffer size, is a configurable parameter. The most fine-grained would be to let the \gls{dl} run for each packet that is received. However, this would be a computational waste and thus, by default, we only let the \gls{dl} run for every 10th packet. Other sampling intervals can be configured. Also, an option would be to not let the \gls{dl} run every $n$th packet but every $x$ milliseconds.

@@

The training procedure involves the following steps:
\begin{enumerate}
\item Draw a random sample of 20 bandwidths, delays, \gls{cca} and flow durations.
\item Simulate each flow concurrently, compute the feature vector continuously and let the neural network output its optimal buffer size each time a new packet arrives. 
\item During each flow, at a random time between 0 and half the flow length in seconds, perform the experiment: Fork the simulation process for each flow and continue one simulation with the current buffer size $+1$ packet and one with the current buffer size $-1$ packet until the end of the flow. This procedure is essentially \textit{A/B testing}.  
\item Wait until all flows are finished
\item Check for each flow which of the two versions performed better ($+1$ or $-1$) regarding the reward function.
\item Update the neural network so that it learns to output the better buffer size when being fed the inputs as they were at the time at which the experimentation started. Specifically, we compute the \gls{mae} between the output of the neural network and the desired output (the one that performed better in the experiment) for the whole batch of 20 results.
\item Start the next iteration. 
\end{enumerate}

@@

When performing offline learning with the tradeoff parameter $\alpha=0.01$ (\autoref{fig:offlineTraining}), the neural network first starts with a prediction that is close to 0 (due to the initialization of neural network weights). During the first couple of thousand training flows, it increases its output and approaches what appears to be the buffer size that works best on average (slightly higher than 15). Then the neural network gradually starts understanding and differentiating different flows and learns custom policies for them. After around 250000 flows, the output appears to be stable and doesn't change anymore. There's one noticeable outlier at around 120000 flows. This appears to be some instability that can commonly occur in the early stages of neural network training but which has no significance for the training process since it resumes normally after the outlier.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_0_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} (offline training, $\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoLargeBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} (offline training, $\alpha=0.01$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleRenoSmallBw}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-6-10-24-18_81000.weights_cc_1_bw_25.000000_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a BIC flow controlled by \gls{ours} (offline training, $\alpha=0.01$) at the bottleneck with a 25\,Mbit/s link and a delay of 15\,ms. The queue never becomes empty and also never forms a standing queue.}
\label{fig:exampleBicLargeBw}
\end{figure}

@@

\autoref{tab:avgOfflineSmallAlpha} shows that the average output maximum queue size is more than double for New Reno than for BIC. This shows that the neural network also learned to distinguish between these two \glspl{cca}. Since there are no features that directly indicate the \gls{cca}, the neural network must have learned these by combining the other features and observing distinct behavior for New Reno and BIC. This is expected, since BIC has a different multiplicative decrease factor than New Reno ($0.7$ vs $0.5$) the neural network should learn to output different maximum queue sizes for these two. Specifically, we would expect BIC to need a smaller buffer as its decrease is smaller, thus requiring a smaller buffer.

\begin{figure*}[h]
\centering
\subfloat[New Reno, varying bandwidth.\label{fig:SmallAlphaNewRenoBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_New_Reno_bandwidth"}.pdf}
}{}
\subfloat[New Reno, varying delay.\label{fig:SmallAlphaNewRenoDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_New_Reno_delay"}.pdf}
}{}
\subfloat[BIC, varying bandwidth\label{fig:SmallAlphaBicBandwidth}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_Bic_bandwidth"}.pdf}
}{}
\subfloat[BIC, varying delay\label{fig:SmallAlphaBicDelay}
]{
\includegraphics[width=0.98\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_logs_plots_2020-6-6-10-24-18_81000.weights_Bic_delay"}.pdf}
}{}
\caption{Change of the buffer size after offline learning when varying bandwidth/delay/\gls{cca}. When bandwidth is varied, delay is kept in the middle of its range, at 15\,ms. Vice versa, when delay is varied, bandwidth is kept in the middle of its range, at 15\,Mbit/s.}
\label{fig:offlineSmallAlpha}
\end{figure*}

@@

\autoref{fig:offlineTrainingLargeAlpha} shows that also for the larger $\alpha$, the neural network converges quickly after a couple of 10000 flows.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{{"figures/local_ns-allinone-3.30.1_ns-3.30.1_results_RLQueueDisc_queueTraces_2020-6-4-15-39-25_108000.weights_cc_0_bw_6.010101_delay_15.000000"}.pdf}
\caption{Showing the maximum queue length and queue length of a New Reno flow controlled by \gls{ours} (offline training, $\alpha=10$) at the bottleneck with a 6\,Mbit/s link and a delay of 15\,ms. The queue occasionally becomes empty which is intended as with the higher tradeoff $\alpha$, \gls{ours} tries to favor a small queue more than high throughput.}
\label{fig:exampleRenoSmallBwLargeAlpha}
\end{figure}

@@

Results in \autoref{tab:corrOfflineLargeAlpha} and \autoref{tab:corrOfflineLargeAlpha} show that for a larger $\alpha$ \gls{ours} still works as expected and that the average output maximum queue size is smaller, as expected.

@@

In contrast to offline learning, online learning enables to train in deployment. This makes it possible to adapt the behavior slowly over time (over a time span of month or years), for example if new \glspl{cca} emerge. Furthermore, being able to training with real flows is more realistic than training only with simulated flows in a simulation.

@@

We envision \gls{ours} being used on switches and routers close to the Internet's edge, as here usually the bottleneck links are and because network devices here only have to manage a rather small number of flows compared to the Internet's core. An interesting future direction of work could be to design and evaluate different reward functions, for example directly based on \gls{qoe}.
